---
title: "Final Project Presentation"
date: "December 9, 2020"
author: "Lauren Walko"
output: html_notebook
---

# Guiding Question

Does mask wearing have an effect on the amount of positive COVID-19 cases in a State?

## Project Background 

The goal of my project is to determine if mask wearing has an effect on the number of positive COVID-19 cases in a state. I found three data sets to use. I have found the states with the highest number of cases, and I am going to use that to compare to the mask wearing frequency in those states. 

## Biggest Challenge

The biggest challenge with my project has been matching the zip codes in the mask data  to determine what state they are from.
I have overcome it by finding a data set to use that I can match the zip codes with.

## Load packages

```{r}
library(DataComputing)
library(tidyverse)
library(mosaic)
library(lubridate)
```

## Load the Data

This data set was a survery done by the New York Times regarding mask wearing in an area. 

```{r}
masks <- read.csv("/home/ltw5073/Final Project/masks.csv")
head(masks)
```

```{r}
cases <- read.csv("/home/ltw5073/Final Project/cases.csv")
head(cases)
```

```{r}
data("ZipGeography")
head(ZipGeography)
```

```{r}
caseperstate <-
  cases %>%
  group_by(state) %>%
  summarise(total = sum(daily.increase))
```

```{r}
caseperstate %>%
  arrange(desc(total))
```

## Bar plot of states with more than 400,000 cases

```{r}
caseperstate %>%
  filter(total > 400000) %>%
  ggplot(aes(x=state, y=total)) +
  geom_bar(stat = 'identity',position='dodge',width=.9)
```

## Final Thoughts

My motivation for this project is to show whether the use of masks has helped a state have less positive cases. It has been an issue that some states wear masks, while others do not, so I want to determine if the states that have consisently worn them have less cases. 