---
title: "Final Project"
author: "Lauren Walko"
date: "November 13, 2020"
output: html_notebook
---


# Guiding Question

## Does mask wearing have an effect on the amount of positive COVID-19 cases in a State?



### Load packages
```{r}
library(DataComputing)
library(tidyverse)
library(mosaic)
library(ggplot2)
```

### Load and inspect the data

#### Mask data
This data came from a survey regarding mask wearing in an area. This data will be used to determine the average responses for each state. 
```{r}
masks <- read.csv("/home/ltw5073/Final Project/masks.csv")
head(masks)
```

#### Cases data
This data comes shows the increase in cases each day for each state.
```{r}
cases <- read.csv("/home/ltw5073/Final Project/cases.csv")
head(cases)
```

#### County Fips data 
This data set will be used to match the zip codes with each state to help determine how each state responded.
```{r}
fips <- read.csv("/home/ltw5073/Final Project/countyfips.csv")
glimpse(fips)
```

#### ZipGeography data
This data set is used to find population for each state.
```{r}
data("ZipGeography")
nrow(ZipGeography)
```

#### States data
I will need this later in my project to join the data sets together with one being abbreviations and the other being the full state name.
```{r}
stabr <- read.csv("/home/ltw5073/Final Project/States.csv")
str(stabr)
```


### Match the fip codes in the mask data to zip codes in the county fips data

```{r}
ZipCodes <- 
  masks %>%
  select(COUNTYFP, NEVER, RARELY, SOMETIMES, FREQUENTLY, ALWAYS) %>%
  left_join(fips %>% select(ZIP, STATE, STCOUNTYFP),
            by = c("COUNTYFP" = "STCOUNTYFP"))
```


### Sort the mask data by state and find the average for each response for each state

```{r}
ZipCodes$STATE <- forcats::fct_explicit_na(ZipCodes$STATE) #Removes missing data from dataset
States <-
  ZipCodes %>%
  group_by(STATE) %>%
  summarise(Never = mean(NEVER), 
            Rarely = mean(RARELY), 
            Sometimes = mean(SOMETIMES),
            Frequently = mean(FREQUENTLY),
            Always = mean(ALWAYS))
```

### Match the state abbreviation with full name of state for later joining 

```{r}
States <- 
  States %>%
  select(Never, Rarely, Sometimes, Frequently, Always, STATE) %>%
  left_join(stabr %>% select(State, Abbreviation),
            by = c("STATE" = "Abbreviation"))
```

#### Show the mask wearing frequencies in descending order by "Never" response

```{r}
States %>%
  arrange(desc(Never))
```

### Determine the cumulative amount of cases that each state has had

```{r}
caseperstate <-
  cases %>%
  group_by(state) %>%
  summarise(cases = sum(daily.increase))
```

#### Arrange the cumulative cases in descending order

```{r}
caseperstate %>%
  arrange(desc(cases))
```

#### Create a bar graph with states that have more than 400,000 cases. 

```{r}
caseperstate %>%
  filter(cases > 400000) %>%
  ggplot(aes(x=state, y=cases)) +
  geom_bar(stat = 'identity',position='dodge',width=.9)
```

This graph shows that California, Florida, Georgia, Illinois, Michigan, New York, Ohio, Pennsylvania, Texas, and Wisconsin have the highest cumulative amount of cases. However, population is an important factor, so I am going to look at the population for each of the states.

### Determine population per state and determine the percent of positive cases per state 

#### Find state population

```{r}
ZipGeography$State <- forcats::fct_explicit_na(ZipGeography$State) #Removes missing data from dataset
Population <-
  ZipGeography %>%
  group_by(State) %>%
  summarise(Pop = sum(Population, na.rm=TRUE))

Population %>%
  arrange(desc(Pop))
```

I am not surprised by what I have found, as the states with the highest population are what was found to have the highest cumulative number of cases.

### Find percent positive for each state

A more effective way to look at the data to compare with the mask wearing data set is to find the percent positive per state.

```{r}
PercentPositive <-
  Population %>%
  select(State, Pop) %>%
  left_join(caseperstate %>% select(state, cases),
            by = c("State" = "state")) %>%
  mutate(PositivePercent = (cases / Pop)*100)
```

#### Arrange the percent positive for each state in descending order

```{r}
PercentPositive %>%
  arrange(desc(PositivePercent))
```

#### Plot the positive percents with states that have a positive percent above 8.00%

```{r}
PercentPositive %>%
  filter(PositivePercent > 8) %>%
  ggplot(aes(x=State,y=PositivePercent)) +
  geom_bar(stat = 'identity',position='dodge',width=.9)
```

### Plot the percent positive cases versus the Never result from survey

```{r}
Covid <-
  PercentPositive %>%
  select(State, PositivePercent) %>%
  left_join(States %>% select(Never, State),
            by = c("State" = "State"))
```

```{r}
Covid %>%
  filter(Never > 0.05) %>%
  ggplot(aes(x=PositivePercent, y=Never, fill=State)) +
  geom_bar(stat = 'identity', position = 'dodge', width = .2)
```

This graph shows an interesting pattern because it appears a high positive percent has a higher "Never" response, but I want to look at the linear regression of the data to make my final determinations. 

### Find the correlation between positive percent and the "Never" response 

```{r}
Covid %>%
  ggplot(aes(x=PositivePercent, y=Never)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```
### Find r^2 value

```{r}
fit1 <- lm(Covid$PositivePercent ~ Covid$Never)
summary(fit1)
```

## Conclusion 

This graph shows that there is a positive relationship between states that were more likely to respond to never wearing a mask and states with a high positive percent. With a r^2 value of 0.4253, there is a semi-strong, positive correlation between the positive percentage in a state and them responding that they never wear a mask. Although, correlation does not equal causation, and while the data shows masks work, other factors are important in stopping the spread of COVID-19.
